\Title

### 2021-08-27 latex/paper/main.tex
Full editing pass of the proof. It is still way too long and needs modifications because some parts feel wrong.

\subsection{All acrobots will rotate}

@@

Using this fact, we now prove that almost every orbit in \(\mathcal{O}_1\) will
cross the line \(|q_u| = \pi\).
Let \(x(0) \in \mathcal{O}_1\) and suppose that the orbit
\(x([0,\infty[)\) does not hit the line \(|q_u| = \pi\).
We claim that this orbit must remain in the compact set 
\(\Sone \times [-K,K]\), for some \(K > 0\).
Suppose otherwise;
then for arbitrary \(K > 0\), there exists some \(t_K > 0\) such that
\(|p_u(t_K)| = K\). 
Looking at the constrained dynamics \eqref{eqn:acrobot-constrained-dynamics},
we see that \(\dot{p}_u \geq -3mgl\) and
\(\dot{q}_u \geq c_1 p_u + c_2\) for some constants \(c_1,c_2 > 0\).
This implies there exists a \(t_{1/2} > 0\) (independent of \(K\)) 
where
\(|p_u(t)| >= K/2\) 
and \(\dot{q}_u \geq c_1 K/2 + c_2\) 
for \(t \in [t_K, t_K + t_{1/2}]\).
Picking \(K\) large enough guarantees \(q_u(t)\) covers a distance of \(2\pi\)
over this time interval, so the orbit would hit the line \(|q_u| = \pi\).

If the orbit \(x(t)\) does not hit the line \(|q_u| = \pi\), we now know it must
remain within the compact set \(\Sone \times [-K,K]\).
By the Birkhoff theorem, \(x(t)\) approaches a positive limit set \(L^+\), 
which is either the upright equilibrium \(\{(\pi,0)\}\), or
a closed orbit not fully contained in \(\mathcal{O}_1\)
which crosses the positive \(q_u\)-axis at some point \(\bar{q} < \pi\).
Suppose \(L^+\) is the latter.
Since \(]0,\mu]\) is negatively invariant for \eqref{eqn:r-discrete}, 
we must have \(\bar{q} > \mu\).
The Poincar\'{e} map \eqref{eqn:poincare-map} is smooth on \(]0,\pi[\), and its
image contains \(\bar{q}\), so \(P^{-1}(\bar{q}) \neq \emptyset\).
Take \(q^\prime \in P^{-1}(\bar{q})\). 
If \(q^\prime \in \, ]0,\mu]\), then \(q^\prime < \bar{q}\);
if \(q^\prime > \mu\), then we know 
from the proof of Lemma \ref{lemma:energy-regulation-D} that
\(P(q^\prime) = \bar{q} \geq q^\prime + \gamma\) for some \(\gamma > 0\), 
so \(q^\prime \leq \bar{q} - \gamma < \bar{q}\).
In either case, \(q^\prime < \bar{q}\), so the orbit starting at \(q^\prime\)
stays within \(\mathcal{O}_1\) and intersects the closed orbit at \(\bar{q}\).
This violates uniqueness of solutions for the constrained dynamics,
so \(L^+\) must be \(\{(\pi,0)\}\), which in turn implies that 
\(x(0) \in \Pi^+\).

>>>>

\subsection{All Acrobots Rotate}

@@

Using this fact, we now prove that almost every orbit in \(\mathcal{O}_1\) will
cross the line \(|q_u| = \pi\).
Let \(x(0) \in \mathcal{O}_1\) and suppose that the orbit
\(x(t)\) does not hit the line \(|q_u| = \pi\).
We claim that this orbit must remain in the compact set 
\(\Sone \times [-K,K]\), for some \(K > 0\).
Suppose otherwise;
then for arbitrary \(K > 0\), there exists some \(t_K > 0\) such that
\(|p_u(t_K)| = K\). 
From \eqref{eqn:acrobot-constrained-dynamics},
we see that \(-3mgl \leq \dot{p}_u \leq 3mgl\) and
\(\dot{q}_u \geq c_1 p_u - c_2\) for some constants \(c_1,c_2 > 0\).
This implies there exists a \(t_{1/2} > 0\) (independent of \(K\)) 
where
\(|p_u(t)| >= K/2\) 
and \(|\dot{q}_u(t)| \geq c_1 K/2 - c_2\) 
for \(t \in [t_K, t_K + t_{1/2}]\).
Picking \(K\) large enough guarantees \(q_u(t)\) covers a distance of \(2\pi\)
over this time interval, so the orbit would hit the line \(|q_u| = \pi\).

We now know that if \(x(t)\) does not hit the line \(|q_u| = \pi\),
it must remain within the compact set \(\Sone \times [-K,K]\).
By the Birkhoff theorem, \(x(t)\) approaches a positive limit set \(L^+\), 
which is either the upright equilibrium \(\{(\pi,0)\}\), or
a closed orbit (not fully contained in \(\mathcal{O}_1\))
which crosses the positive \(q_u\)-axis at some point \(\bar{q} < \pi\).
Suppose \(L^+\) is the latter.
Since \(]0,\mu]\) is negatively invariant for \eqref{eqn:r-discrete}, 
we must have \(\bar{q} > \mu\).
The Poincar\'{e} map \eqref{eqn:poincare-map} is smooth on \(]0,\pi[\), and its
image contains \(\bar{q}\), so \(P^{-1}(\bar{q}) \neq \emptyset\).
Take \(q^\prime \in P^{-1}(\bar{q})\). 
If \(q^\prime \in \, ]0,\mu]\), then \(q^\prime < \bar{q}\);
if \(q^\prime > \mu\), then we know 
from the proof of Lemma \ref{lemma:energy-regulation-D} that
\(P(q^\prime) = \bar{q} \geq q^\prime + \gamma\) for some \(\gamma > 0\), 
so \(q^\prime \leq \bar{q} - \gamma < \bar{q}\).
In either case, \(q^\prime < \bar{q}\), so the orbit starting at \(q^\prime\)
stays within \(\mathcal{O}_1\) and intersects the closed orbit at \(\bar{q}\).
This violates uniqueness of solutions for the constrained dynamics,
so \(L^+\) must be \(\{(\pi,0)\}\), which in turn implies that 
\(x(0) \in \Pi^+\).
